---
title: 'Digital Attribution at W.M. Winters'
author: "Adil Wahab"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
  word_document: default
---

```{r message=TRUE, warning=TRUE, include=FALSE}
library(tidyverse)
library(data.table)
library(ggplot2)
library(janitor)
library(reshape2)
library(kableExtra)
load("/Users/Adil/Desktop/Data Science Projects/R/Digital Marketing/Winters_Attribution.rdata")
winters = data
```

# Project Background
The project uses the “Winters_Attribution” data. The Winters case study uses a unique data set was collected in collaboration with a large online media analytics and optimization platform company. The online media company managed the entire campaign of a U.S.-based retailer. The individual-level data set consists of advertising exposures and user-initiated actions, with users tracked across different advertising channels and media. Note that all observations relate to an order (touchpoints that do not lead to a purchase are absent). The unit of observation is an order-touchpoint, so that the same order is repeated by the number of touches. A brief description of the variables follows:
1. Orderid: actual transaction id from the vendor — if there are multiple touch points, you will see this ID listed on each row
2. Orderdatetime: UTC time of the transaction
3. Saleamount: value of the transaction
4. Newcustomer: Y = yes // N = no
5. Position: touchpoint position
6. Positiondatetime: UTC time of the actual touchpoint
7. Groupname: the group/channel by which the client categorized their marketing activities (e.g. CPM = display advertising, BUZZ AFFILIATE & CJ = affiliates)
8. Networkname: this is the name of the media touchpoint
9. Networkid: this is the id used by C3
10. Brand: this is used for search terms, where N = 'Non-Brand' and Y = 'Brand Search'
11. Positionname: C3 Metrics terminology for the touchpoints.
a) “Originator”: the first touchpoint
b) “Converter”: the last touchpoint
c) For more than 2 touchpoints:
i. “Assist”: the penultimate touchpoint
ii. “Roster”: any touchpoints between “Originator” and “Assist”
12. DaysToConvert: Days between first touch and conversion
13. Touches: Number of touchpoints

# EDA
```{r}
head(winters,5)
```

# 1. First-touch vs Last-touch
## a.
```{r}
# create new columns for originator and converter
winters$Converter = ifelse(winters$Positionname == "CONVERTER", 1, 0)
winters$Originator = ifelse(winters$Positionname == "ORIGINATOR", 1, 0)
n = length(unique(winters$Orderid))
# create table with new columns based on media channels 
winters %>% group_by(Groupname) %>%
  summarise(converter = sum(Converter), originator = sum(Originator)) %>%
  mutate(converter_per = round(converter / n,4) * 100,
         originator_per = round(originator/ n,4) * 100) %>% 
  select(Groupname, converter, originator, converter_per, originator_per) -> d
# add row of column totals 
d[14,] = c("Total", colSums(d[,2:5]))
kable(d)
```

## b. Share of credit for the first- and last-touch attribution models
```{r message=FALSE, warning=FALSE}
d %>% select(Groupname, converter_per, originator_per) -> d1
d1 <- d1[1:13,]
dfa <- melt(d1, id.vars="Groupname");

ggplot(dfa) + geom_col(aes(x=Groupname, y=as.numeric(value),
                                       fill = variable), position='dodge') +
  labs(y="Share of Credit (%)", x="Marketing Channel", fill = "Attribution Model") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(labels = c("Last-touch", "First-touch"),values = c("blue", "red"))
```

## c. Interpretation
There appears to be a wider distribution spread among the marketing channels for the first-touch attribution model compared to the last-touch model. The last-touch model is dominated by Buzz Affiliate, CJ, and CPM, whereas, the first-touch model's main channels are CPM and branded search on Google. 

Allocating marketing budget only on the basis of the last-touch attribution model doesn't account for other prominent marketing channels a customer interacted with. This attribution model can be detrimental for Winters as it will not take budget on search-based ads, particularly branded search on Google, into consideration. Following a last-touch model for marketing budget allocation means a very concentrated investment in Buzz Affiliate, CJ and CPM. This attribution model would miss out on the diversified investment in more marketing channels that a first-touch allocation approach would take. 

# 2. New Customers vs. Old Customers
## a. Days to convert
```{r}
winters %>% filter(Originator == 1) -> org
org %>% filter(Newcustomer == "Y") %>% summarise(mean = mean(DaysToConvert)) -> newC
org %>% filter(Newcustomer == "N") %>% summarise(mean = mean(DaysToConvert)) -> oldC

print(paste0("Average number of days for a new customer to convert: ", round(newC,0), " days"))

print(paste0("Average number of days for an old customer to convert: ", round(oldC,0), " days"))
```

## b. Average touchpoints 
```{r}
winters %>% group_by(Orderid) %>%
  select(Orderid, Position, Originator, Newcustomer) %>% add_count() %>% filter(Originator == 1, Newcustomer == "Y") -> tp_n
winters %>% group_by(Orderid) %>%
  select(Orderid, Position, Originator, Newcustomer) %>% add_count() %>% filter(Originator == 1, Newcustomer == "N") -> tp_o

avg_tp_n = round(mean(tp_n$n),2)
avg_tp_o = round(mean(tp_o$n),2)

print(paste0("Average number of touchpoints for a new customer: ", avg_tp_n, " touches"))

print(paste0("Average number of touchpoints for an old customer: ", avg_tp_o, " touches"))
```

## c. Average order sales
```{r}
org %>% filter(Newcustomer == "Y") %>% summarise(mean = mean(Saleamount)) -> O_newC
org %>% filter(Newcustomer == "N") %>% summarise(mean = mean(Saleamount)) -> O_oldC

print(paste0("Average order sales for new customers: $", round(O_newC,2)))

print(paste0("Average order sales for old customers: $", round(O_oldC,2)))
```

## d. Summary of the three variables
New customers have a lower average touchpoint count and they take significantly fewer days to convert. The average sales for a new customer is also higher than an old customer by more than $60. 

# 3. Revenue per marketing channel based on first-touch attribution
## a. Average sales and total revenue
```{r}
winters %>% group_by(Groupname) %>% filter(Positionname == "ORIGINATOR") %>% 
  summarise(avg_sales = round(mean(Saleamount),2), revenue = sum(Saleamount)) %>% 
  select(Groupname, avg_sales, revenue) -> avg_r
kable(avg_r)
```

## b. Total incremental gross revenue
```{r message=FALSE, warning=FALSE}
# gross margin = 40%, incr_fact_brand = 5%, incr_fact_rem = 10%

winters %>% group_by(Groupname, Brand) %>%
  filter(Positionname == "ORIGINATOR") %>% 
  mutate(gross_rev = ifelse(Brand == "Y", Saleamount*0.4*0.05, Saleamount*0.4*0.1)) %>% 
  summarise(gross_margin = round(sum(gross_rev),2)) -> b3

b3 %>% filter(Brand == "Y") -> branded
branded[5,] = c("Total","NA", colSums(branded[,3]))

kable(b3)
kable(branded)
print(paste0("The total incremental gross revenue accuring to Winters by originator channel of branded search is $", branded[5,3]))
```
The total incremental gross revenue accuring to Winters by originator channel of branded search is $2,945.71

## c. Advice for the search team
According to the calculation above, the branded search adversiting cost is significantly higher than the incremental gross revenue (by first-touch attribution). The cost is $4,200 while the revenue is only $2,945.71. The search ad team should avoid spending this much on branded search ads as it is exceeding the revenue by a significant amount. However, it is important to not fully neglect branded search ads as we need to keep up with our competitors. 

# 4. Linear/uniform attribution
## a. 
```{r}
winters %>% group_by(Orderid) %>%
  select(Orderid, Position, Originator, Newcustomer, Groupname) %>% add_count() %>%  mutate(LinearAttributionShare = 1/n) -> a4

a4 %>% group_by(Groupname) %>% 
  summarise(total_share = round(sum(LinearAttributionShare),2)) %>% arrange(desc(total_share)) -> done
done %>% mutate(credit_share = round(as.numeric(total_share)/1518,4)*100) -> k
k[14,] = c("Total", colSums(k[,2:3]))
kable(k)
```

## b. Plot of all 3 attribution models
```{r}
df2 <- k[1:13,]
df2 %>% select(Groupname, credit_share) -> df2
dfb <- melt(df2, id.vars="Groupname")
comb <- rbind(dfa,dfb)

ggplot(comb) + geom_col(aes(x=Groupname, y=as.numeric(value),
                                       fill = variable), position='dodge') +
  labs(y="Share of Credit (%)", x="Marketing Channel", fill = "Attribution Model") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(labels = c("Last-touch", "First-touch","Linear/Uniform"),values = c("blue", "red","green"))
```

## c. Interpretation of the different attribution models
The linear and first-touch models have a more spread out distribution of credit to different marketing channels compared to the last-touch model. However, the linear model appears to share similarities with both of the other models. The last-touch and linear models both give the most credit Buss Affiliate and CPM. On the other hand, the linear and first-touch models both give some credit to branded search ads while the last-touch model gives none. While each one of the three attribution models have their differences, there is no denying of the dominant share of CPM. 

# 5. Intermediate (Roster and Assist) touch points 
## a.
```{r}
prop <- as.data.frame(with(winters, table(Groupname, Positionname)) %>% prop.table(margin=1)*100) %>% filter(Groupname %in% c("BUZZ AFFILIATE","CJ","CPM","SEARCH GOOGLE BRAND","SEARCH GOOGLE NON-BRAND","SEARCH MSN BRAND","TV"))

prop %>% pivot_wider(names_from = Positionname, values_from = Freq) %>% 
  arrange(Groupname) %>% select(Groupname, ORIGINATOR, ROSTER, ASSIST, CONVERTER) %>%
  mutate(Total = (ORIGINATOR+ROSTER+ASSIST+CONVERTER)) -> a5
kable(a5)
```

## b. 
```{r}
ggplot(prop) + geom_col(aes(x=Groupname, y=as.numeric(Freq),
                                       fill = Positionname), position='dodge') +
  labs(y="Share (%)", x="Marketing Channel", fill = "Touchpoint Type") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## c. 
CPM, which is display adversiting, has the highest proportion of roster touchpoints relative to all its other touchpoint types, while Google's non-branded search has the lowest proportion of touchpoints dedicated to the roster. CPM and Buzz Affiliate have the most assist touchpoints compared to all other channels. If you look at assist and roster together, CPM and Buzz Affiliate again seems to be getting the most credit from these two types of touchpoints.

The branded search ads receive zero credit under the last-touch model, which may be an issue since branded search ads can play a very crucial role in intent and awareness where the customers search with the sole purpose of making the purchase or gaining the extra confidence to make the purchase. CPM has an extremely high share of roster credit while getting very low credit from last-touch attribution. This may be an issue since display ads play a fundamental role in creating brand awareness that may just be the bridge to get to a roster touchpoint or the difference-maker between an assist and a conversion. From a first-touch perspective CPM, CJ, and Buzz Affiliate all seem to be receiving relatively low credit compared to all the search and TV ads, which may be misleading since these marketing channels can be influential forces in raising brand awareness and exposure. 
